From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-09 Pi+/paper/discussion.tex
working on sec 8

%%%%%%%%%%%%%%
\paragraph{Rewriting} We presented a rewriting system for the Coxeter relations for $\Sn$ to solve its word problem.

There exists an algorithm, due to~\citet{knuthSimpleWordProblems1970}, that, when succeeds, constructs a well-behaved
rewriting system for an arbitrary finite set of (undirected) equations. It did not work for us, producing too many
equations, and proving correctness and termination was intractable.

We chose to encode permutations as adjacent transpositions corresponding to words in $\Sn$, since that closely
corresponds to the syntax of $\PiLang$ combinators. Other representations of permutations would give a different
denotational semantics for $\PiLang$ combinators. Permutations can be encoded as listed vectors or matrices, inductively
generated trees (Motzkin trees), Young diagrams, or String diagrams, but the difficulty of formalising them in type
theory varies depending on the encoding.

Rewriting systems and word problems have a long history of being formalised in proof assistants. In the recent past,
higher order rewriting systems have been formalised in proof assistants like {homotopy.io} and Lean and Coq. The use of
HoTT to study rewriting has been considered in~\cite{krausCoherenceWellFoundednessTaming2020}.

>>>>

%%%%%%%%%%%%%%
\paragraph{Rewriting} Term-rewriting systems and word problems have a long history of being formalised in proof
assistants. In the recent past, higher order rewriting systems have been formalised in {homotopy.io}, Lean, Coq, and
HoTT~\citet{krausCoherenceWellFoundednessTaming2020}. As part of the proof of our main result, we developed a rewriting
system for the Coxeter relations for $\Sn$ to solve its word problem and formalised its correctness and termination in a
proof assistant. Our design based on encoding permutations as adjacent transpositions led to an intuitive whose
correctness was straightforward to establish and whose termination only required slightly modified variations on
Church-Rosser theorems. Other encodings of permutations as listed vectors, matrices, inductively generated trees
(Motzkin trees), Young diagrams, or string diagrams, proved either difficult formalise in type theory or difficult to
relate directly to the primitive type isomorphisms of $\PiLang$. The automatic \citet{knuthSimpleWordProblems1970}
produced too many equations making proving correctness and termination intractable.
